---
title: ""
author: "Lucy King"
date: "4/17/2020"
output: html_document
---

# Environment
```{r}
library(tidyverse)
library(janitor)
library(lme4)
library(ggsci)
library(ggpubr)
library(lmerTest)
library(performance)
library(effectsize)
library(parameters)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
source("winsorize.R")

# Files
home <- "~/Box/Mooddata_Coordinating/BABIES/Data/final_scored_data/"

hcc_data_file <- paste0(home, "hair_cortisol/hcc_mom_infant_cleaned_all.csv")
demo_file <- paste0(home, "demographics/demo_longitudinal_cleaned_final.csv")
prams_file <- paste0(home, "PRAMS/prams_alldata_cleaned.csv")
cesd_file <- paste0(home, "CESD/cesd_wf_20201001.csv")
crisys_file <- paste0(home, "CRISYS/crisys_scored_wf_complete.csv")

source("R_rainclouds.R")

theme_hcc <-
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 18, hjust = .5),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 16),
    legend.title = element_text(size = 18), 
    legend.text = element_text(size = 16),
    legend.position = "bottom"
  )

cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

options(scipen = 999)

```

# Read in data
```{r}
d <-
  read_csv(hcc_data_file) %>% 
  #retain only participants from prenatal cohort
  filter(ID >= 1000) %>% 
  left_join(
    read_csv(demo_file) %>% 
      select(
        ID,
        education_t1,
        education_t1_txt,
        mom_race,
        mom_latinx,
        annual_income_t1,
        annual_income_t1_txt,
        mom_age_t1,
        income_needs_t1,
        infant_age_t2 = age_scan_t2,
        infant_age_t3 = age_scan_t3
      ),
    by = "ID"
  ) %>% 
  left_join(
    read_csv(prams_file) %>% 
      select(
        ID,
        preg_previous,
        prev_preg_total,
        preg_weight_chg,
        labor_status,
        labor_mode,
        mom_preg_bmi,
        mom_newborn_bmi,
        preg_weight_chg,
        preg_weight_gain,
        preg_weight_loss,
        n_prenatal_cigarettes,
        n_newborn_cigarettes,
        n_prenatal_drinks,
        n_newborn_drinks,
        birth_weight_total_lbs
      ),
    by = "ID"
  )  %>% 
  left_join(
    read_csv(cesd_file) %>% 
      dplyr::select(
        ID,
        cesd_t1,
        cesd_t2,
        cesd_t3
      ), 
    by = "ID"
  ) %>% 
  left_join(
    read_csv(crisys_file) %>% 
      dplyr::select(
        ID,
        crisys_total_T1,
        crisys_total_T2,
        crisys_total_T3
      ), 
    by = "ID"
  ) %>% 
  mutate(
    # recale segment so bottom is 0
    segment = recode(
      segment,
      "1" = 0,
      "2" = 1,
      "3" = 2, 
      "4" = 3,
      "5" = 4
    ),
    season = as.factor(season),
    hair_chemexp_past5mo = factor(hair_chemexp_past5mo),
    phase = factor(
      phase,
      levels = c(
        "1st trimester",
        "2nd trimester",
        "3rd trimester",
        "3mo. postpartum",
        "6mo. postpartum"
      ),
      labels = c(
        "1st trimester",
        "2nd trimester",
        "3rd trimester",
        "3 mo. postpartum",
        "6 mo. postpartum"
      )
    ),
    cesd = case_when(
      timepoint == "T1" ~ cesd_t1,
      timepoint == "T2" ~ cesd_t2,
      timepoint == "T3" ~ cesd_t3
    ),
    crisys = case_when(
      timepoint == "T1" ~ crisys_total_T1,
      timepoint == "T2" ~ crisys_total_T2,
      timepoint == "T3" ~ crisys_total_T3,
    )
  ) %>% 
  group_by(ID) %>% 
  mutate(
    # mean cesd for each participant across timepoints
    cesd_mu = mean(cesd, na.rm = TRUE), 
     # person-mean-centered cesd (deviations from mean)
    cesd_pc = if_else(
      !is.na(cesd), cesd - cesd_mu, NA_real_
    ),
    crisys_mu = mean(crisys, na.rm = TRUE),
    crisys_pc = if_else(
      !is.na(crisys), crisys - crisys_mu, NA_real_
    )
  ) %>% 
  ungroup()
```

# Descriptive statistics


## Total number of participants at each time point, prior to removing observations contemporaneous with corticosteroid use
```{r}
d %>% 
  filter(!is.na(hcc_log_m)) %>% 
  distinct(ID, timepoint) %>% 
  count(timepoint)
```


## Remove maternal observations contemporaneous with corticosteroid use
Four participants did not provide information about recent corticosteroid use
```{r}
# number of participants
d %>% 
  distinct(ID, timepoint, steroid_recent) %>% 
  count(timepoint, steroid_recent)

# number of observations
d %>% 
  count(timepoint, steroid_recent)
```
```{r}
d <-
  d %>% 
  mutate(
    hcc_m = case_when(
      steroid_recent == 1 ~ NA_real_, 
      is.na(steroid_recent) ~ hcc_m,
      steroid_recent == 0 ~ hcc_m, 
    ),
    hcc_log_m = case_when(
      steroid_recent == 1 ~ NA_real_, 
      is.na(steroid_recent) ~ hcc_log_m,
      steroid_recent == 0 ~ hcc_log_m, 
    )
  ) %>% 
  filter(!is.na(hcc_log_m))
```


## Final total number of participants and observations

### N
```{r}
d %>% 
  distinct(ID) %>% 
  count()
```
 
### n 
```{r}
d %>% 
  count()
```
 
### N by timepoint
```{r}
d %>% 
  distinct(ID, timepoint) %>% 
  count(timepoint)
```

### n by timepoint
```{r}
d %>% 
  count(timepoint)
```

### N across timpoints
```{r}
d %>% 
  distinct(ID, timepoint, hcc_log_m) %>% 
  mutate(
    hcc = !is.na(hcc_log_m)
  ) %>% 
  select(-hcc_log_m) %>% 
  distinct(ID, timepoint, hcc) %>% 
  spread(timepoint, hcc) %>% 
  count(T1, T2, T3)
```
### by phase
```{r}
d %>% 
  distinct(ID, phase) %>% 
  count(phase)
```

```{r}
d %>% 
  filter(!is.na(hcc_log_m)) %>% 
  distinct(ID, phase, hcc_log_m) %>% 
  mutate(
    hcc = !is.na(hcc_log_m)
  ) %>% 
  select(-hcc_log_m) %>% 
  distinct(ID, phase, hcc) %>% 
  spread(phase, hcc) 
```

```{r}
d %>% 
  count(phase)
```
## Distributions

### Weeks since conception
```{r}
d %>% 
  ggplot(aes(seg_week, fill = timepoint)) +
  geom_density(alpha = 1/2) +
  theme_hcc +
  scale_x_continuous(breaks = seq.int(0, 65, 10)) +
  scale_fill_manual(values = cbbPalette) +
  theme(
    legend.position = "right"
  ) +
  labs(
    fill = "Assessment\ntimepoint",
    x = "Weeks since conception of fetus",
    y = "Density of HCC observations"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/hair_cortisol/plots/hcc_timepoint_density.png",
  width = 9,
  height = 7
)
```

### HCC in each phase

```{r}
d %>% 
  ggplot(aes(fct_rev(phase), hcc_log_m)) +
  geom_flat_violin(
    aes(color = fct_rev(phase), fill = fct_rev(phase)),
    position = position_nudge(x = .25, y = 0),
    adjust = 2,
    trim = TRUE
  ) +
  geom_point(
    aes(color = fct_rev(phase), fill = fct_rev(phase)),
    position = position_jitter(width = .15),
    size = 1
  ) +
  geom_boxplot(
    aes(x = as.numeric(fct_rev(phase)) + .25, y = hcc_log_m, group = fct_rev(phase)),
    color = "black",
    alpha = 1/3,
    width = .1,
    outlier.shape = NA
  ) +
  scale_y_continuous(breaks = seq.int(-2, 5, 1)) +
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  theme_hcc +
  theme(
    legend.position = "none",
    plot.caption = element_text(hjust = .5, size = 15)
  ) +
  labs(
    x = NULL,
    y = "log(HCC) pg/mg"
  ) +
  coord_flip()

ggsave(
  "~/Box/lucy_king_files/BABIES/hair_cortisol/plots/hcc_rain_clouds_phase.png",
  width = 8,
  height = 7
)
```

### CES-D scores at each timepoint

```{r}
d %>% 
  ggplot(aes(cesd)) +
  geom_histogram(binwidth = 5) +
  theme_hcc +
  labs(
    fill = "Assessment\ntimepoint",
    x = "CES-D Score"
  ) +
  facet_grid(.~timepoint)
```
### CRISYS scores at each timepoint

```{r}
d %>% 
  ggplot(aes(crisys)) +
  geom_histogram(binwidth = 2) +
  theme_hcc +
  labs(
    fill = "Assessment\ntimepoint",
    x = "CRISYS Score"
  ) +
  facet_grid(.~timepoint)
```
## Internal consistency of CESD and CRISYS
```{r}
r <-
  read_csv(cesd_file) %>% 
  left_join(read_csv(crisys_file), by = "ID")
```
### CESD
```{r}
r %>% 
  dplyr::select(
    contains("cesd")
  ) %>% 
  dplyr::select(
    ends_with("_T1"),
    -cesd_t1
  ) %>% 
  psych::alpha()

r %>% 
  dplyr::select(
    contains("cesd")
  ) %>% 
  dplyr::select(
    ends_with("_T2"),
    -cesd_t2
  ) %>% 
  psych::alpha()

r %>% 
  dplyr::select(
    contains("cesd")
  ) %>% 
  dplyr::select(
    ends_with("_T3"),
    -cesd_t2
  ) %>% 
  psych::alpha()
```

### CESD
```{r}
r %>% 
  dplyr::select(
    contains("crisys")
  ) %>% 
  dplyr::select(
    ends_with("_T1"),
    -crisys_total_T1:-crisys_ses_win_T1,
    # removing items below due to high missingness / 0 variance
    -crisys_69_T1,
    -crisys_70_T1,
    -crisys_71_T1
  ) %>% 
  psych::alpha()

r %>% 
  dplyr::select(
    contains("crisys")
  ) %>% 
  dplyr::select(
    ends_with("_T2"),
    -crisys_total_T2:-crisys_ses_win_T2,
    -crisys_69_T2,
    -crisys_70_T2,
    -crisys_71_T2
  ) %>% 
  psych::alpha()

r %>% 
  dplyr::select(
    contains("crisys")
  ) %>% 
  dplyr::select(
    ends_with("_T3"),
    -crisys_total_T3:-crisys_ses_win_T3,
    -crisys_20_T3,
    -crisys_47_T3,
    -crisys_68_T3,
    -crisys_69_T3,
    -crisys_70_T3,
    -crisys_71_T3
  ) %>% 
  psych::alpha()

```

## Demographics, means, SDs, ranges

### Age
```{r}
d %>% 
  summarise_at(
    vars(
      mom_age_t1, 
      weeks_collect_t1, 
      weeks_collect_t2,
      weeks_collect_t3, 
      infant_age_t2,
      infant_age_t3,
      gestational_weeks_birth
    ),
    funs(mean, sd, min, max), na.rm  = TRUE
  )
```

### HCC
```{r}
d %>% 
  group_by(timepoint) %>% 
  summarise_at(
    vars(hcc_m),
    funs(mean, sd, min, max), na.rm = TRUE
  )

d %>% 
  group_by(phase) %>% 
  summarise_at(
    vars(hcc_m),
    funs(mean, sd, min, max), na.rm = TRUE
  )

d %>% 
  group_by(phase) %>% 
  summarise_at(
    vars(hcc_log_m),
    funs(mean, sd, min, max), na.rm = TRUE
  )
```
### Weeks since conception and since birth
```{r}
d %>% 
  group_by(timepoint) %>% 
  summarise_at(
    vars(seg_week),
    funs(mean, sd, min, max), na.rm = TRUE
  )

d %>% 
  group_by(phase) %>% 
  summarise_at(
    vars(seg_week),
    funs(mean, sd, min, max), na.rm = TRUE
  )

d %>% 
  group_by(phase) %>% 
  summarise_at(
    vars(seg_week_post),
    funs(mean, sd, min, max), na.rm = TRUE
  )
```

### Hair treatments, hair washing
```{r}
d %>% 
  distinct(ID, timepoint, hair_chemexp_past5mo) %>% 
  count(timepoint, hair_chemexp_past5mo)

d %>% 
  distinct(ID, timepoint, hair_wash_freq) %>% 
  count(timepoint, hair_wash_freq)

d %>% 
  distinct(ID, hair_wash_freq_mean) %>% 
  count(is.na(hair_wash_freq_mean))
  
d %>% 
  summarise_at(
    vars(hair_wash_freq_mean),
    funs(mean, sd, min, max), na.rm = TRUE
  )
```

### CES-D scores
```{r}
d %>% 
  distinct(ID, timepoint, cesd) %>% 
  count(timepoint, !is.na(cesd))
```

```{r}
d %>% 
  distinct(ID, timepoint, cesd) %>% 
  group_by(timepoint) %>% 
  summarise_at(
    vars(cesd),
    funs(mean, sd, min, max)
  )
```

```{r}
d %>% 
  distinct(ID, timepoint, cesd) %>% 
  count(timepoint, cesd >=16)
```

### CRISYS scores
```{r}
d %>% 
  distinct(ID, timepoint, crisys) %>% 
  count(timepoint, !is.na(crisys))
```

```{r}
d %>% 
  distinct(ID, timepoint, crisys) %>% 
  group_by(timepoint) %>% 
  summarise_at(
    vars(crisys),
    funs(mean, sd, min, max), na.rm = TRUE
  )
```
### Race/ethnicity
```{r}
d %>% 
  distinct(ID, mom_race) %>% 
  count(mom_race)

d %>% 
  distinct(ID, mom_latinx) %>% 
  count(mom_latinx)
```


### Income 
```{r}
d %>% 
  distinct(ID, annual_income_t1_txt) %>% 
  count(annual_income_t1_txt)

```
# Missing data analyses
```{r}
dm <-
  d %>% 
  dplyr::select(
    ID,
    timepoint,
    mom_age_t1,
    gestational_weeks_birth,
    cesd_t1:cesd_t3,
    crisys_total_T1:crisys_total_T3
  ) %>% 
  distinct(ID, timepoint, .keep_all = TRUE) %>% 
  mutate(
    T2_complete = if_else(timepoint == "T2", TRUE, FALSE),
    T3_complete = if_else(timepoint == "T3", TRUE, FALSE)
  ) %>% 
  group_by(ID) %>% 
  mutate(
    T2_complete = sum(T2_complete),
    T3_complete = sum(T3_complete)
  ) %>% 
  ungroup() %>% 
  distinct(ID, .keep_all = TRUE) 
```

```{r}
t.test(dm$mom_age_t1 ~ dm$T2_complete)
t.test(dm$gestational_weeks_birth ~ dm$T2_complete)
t.test(dm$cesd_t1 ~ dm$T2_complete)
t.test(dm$cesd_t2 ~ dm$T2_complete)
t.test(dm$cesd_t3 ~ dm$T2_complete)
t.test(dm$crisys_total_T1 ~ dm$T2_complete)
t.test(dm$crisys_total_T2 ~ dm$T2_complete)
t.test(dm$crisys_total_T3 ~ dm$T2_complete)
```

```{r}
t.test(dm$mom_age_t1 ~ dm$T3_complete)
t.test(dm$gestational_weeks_birth ~ dm$T3_complete)
t.test(dm$cesd_t1 ~ dm$T3_complete)
t.test(dm$cesd_t2 ~ dm$T3_complete)
t.test(dm$cesd_t3 ~ dm$T3_complete)
t.test(dm$crisys_total_T1 ~ dm$T3_complete)
t.test(dm$crisys_total_T2 ~ dm$T3_complete)
t.test(dm$crisys_total_T3 ~ dm$T3_complete)
```

# Results

## Reliability of segments 4- and 5-cm distal from scalp

```{r}
do5_T1T2 <-
  d %>% 
  select(
    ID, 
    timepoint,
    hcc_m,
    segment,
    seg_week
  ) %>% 
  filter(
    (segment == 0 & timepoint == "T1") |
    (segment == 4 & timepoint == "T2") 
  ) %>% 
  group_by(ID) %>% 
  mutate(
    n = n()
  ) %>% 
  filter(n == 2) %>% 
  arrange(ID, timepoint) %>% 
  group_by(ID) %>% 
  mutate(
    seg_week_diff = seg_week - lag(seg_week)
  ) %>% 
  group_by(ID) %>% 
  mutate(
    seg_week_diff = sum(seg_week_diff, na.rm = TRUE)
  ) %>% 
  filter(seg_week_diff < 4 & seg_week_diff >= -4) %>% 
  select(ID, timepoint, hcc_m) %>% 
  spread(timepoint, hcc_m) %>% 
  rename(
    proximal = T1,
    distal = T2
  ) %>% 
  ungroup()
```

```{r}
do5_T2T3 <-
  d %>% 
  select(
    ID, 
    timepoint,
    hcc_m,
    segment,
    seg_week
  ) %>% 
  filter(
    (segment == 0 & timepoint == "T2") |
    (segment == 4 & timepoint == "T3") 
  ) %>% 
  group_by(ID) %>% 
  mutate(
    n = n()
  ) %>% 
  filter(n == 2) %>% 
  arrange(ID, timepoint) %>% 
  group_by(ID) %>% 
  mutate(
    seg_week_diff = seg_week - lag(seg_week)
  ) %>% 
  group_by(ID) %>% 
  mutate(
    seg_week_diff = sum(seg_week_diff, na.rm = TRUE)
  ) %>% 
  filter(seg_week_diff < 4 & seg_week_diff >= -4) %>% 
  select(ID, timepoint, hcc_m) %>% 
  spread(timepoint, hcc_m) %>% 
  rename(
    proximal = T2,
    distal = T3
  ) %>% 
  ungroup()
```

```{r}
do4_T1T2 <-
  d %>% 
  select(
    ID, 
    timepoint,
    hcc_m,
    segment,
    seg_week
  ) %>% 
  filter(
    (segment == 0 & timepoint == "T1") |
    (segment == 3 & timepoint == "T2") 
  ) %>% 
  group_by(ID) %>% 
  mutate(
    n = n()
  ) %>% 
  filter(n == 2) %>% 
  arrange(ID, timepoint) %>% 
  group_by(ID) %>% 
  mutate(
    seg_week_diff = seg_week - lag(seg_week)
  ) %>% 
  group_by(ID) %>% 
  mutate(
    seg_week_diff = sum(seg_week_diff, na.rm = TRUE)
  ) %>% 
  filter(seg_week_diff < 3 & seg_week_diff >= -3) %>% 
  select(ID, timepoint, hcc_m) %>% 
  spread(timepoint, hcc_m) %>% 
  rename(
    proximal = T1,
    distal = T2
  ) %>% 
  ungroup() 
```

There are no overlapping T2, T3 segments for 3-4cm from scalp.

```{r}
do <-
  do5_T1T2 %>% 
  bind_rows(do5_T2T3) %>% 
  bind_rows(do4_T1T2) %>% 
  mutate(
    proximal_log = log(proximal),
    distal_log = log(distal)
  ) 

do %>% 
  count()

do %>% 
  distinct(ID) %>% 
  count()
```

```{r}
do %>% 
  select(proximal, distal) %>% 
  psych::ICC()

cor.test(do$proximal, do$distal)
t.test(do$proximal, do$distal, paired = TRUE)
```
```{r}
do  %>% 
  ggplot(aes(log(proximal), log(distal))) +
  geom_point(size = 4) +
  geom_smooth(method = "lm", size = 3) +
  theme_hcc +
  labs(
    x = "log(HCC) in proximal hair\n(0-1cm from scalp)",
    y = "log(HCC) in distal hair\n(3-5cm from scalp)"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/hair_cortisol/plots/hcc_distal_validity.png",
  width = 7,
  height = 5
)
```
```{r}
do  %>% 
  select(
    proximal_log:distal_log
  ) %>% 
  gather(
    sample,
    value,
    proximal_log:distal_log
  ) %>% 
  ggplot(aes(sample, log(value))) +
  geom_boxplot() +
  geom_point(alpha = 1/2) +
  theme_hcc +
  labs(
    y = "log(HCC) pg/mg",
    x = "Sample type"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/hair_cortisol/plots/hcc_distal_validity_boxplot.png",
  width = 7,
  height = 5
)
```


```{r}
d3 <-
  d %>% 
  filter(segment < 4)

d3 %>% 
  distinct(ID, timepoint) %>% count(timepoint)
```

## Fit piecewise linear mixed model of HCC, covarying for segment as a random effect

Piecewise model is centered at birth of infant (due date if birth date never provided)
```{r}
mlm1 <- lmer(
  scale(hcc_log_m) ~ 
    scale(seg_week_preg) +
    scale(seg_week_post) +
    (scale(seg_week_preg) + scale(seg_week_post) + scale(segment) | ID),
  data = d
)

summary(mlm1)
mlm1_parameters <- model_parameters(mlm1)
mlm1_parameters
performance(mlm1)
tab_model(mlm1)
```

## Visualize piecewise model

### Extract predicted values and SEs

```{r}
# asymptotic covariance matrix (variances and covariances of fixed effects)
acov_mlm1 <- vcov(mlm1) 

pred_mlm1 <- expand.grid(
  seg_week_cent_birth = seq(-40, 30, 1)
) %>% 
  mutate(
    seg_week_preg = scale(
      if_else(
        seg_week_cent_birth <= 0, seg_week_cent_birth, 0
      )
    ),
    seg_week_post = scale(
      if_else(
        seg_week_cent_birth > 0, seg_week_cent_birth, 0
      )
    )
  ) 

xmat_mlm1 <-
  pred_mlm1 %>% 
  mutate(
    intercept = 1
  ) %>% 
  dplyr::select(
    intercept,
    seg_week_preg,
    seg_week_post
  ) %>% 
  as.matrix()

# predicted values
fixed_mlm1 <- summary(mlm1)$coefficients[,"Estimate"] 

pred_mlm1$fit <- 
  xmat_mlm1 %*% fixed_mlm1

# standard errors
pred_mlm1$se <- 
  xmat_mlm1 %*% acov_mlm1 %*% t(xmat_mlm1) %>% 
  diag() %>% 
  sqrt()

# CIs
pred_mlm1 <- 
  pred_mlm1 %>% 
  mutate(LL = fit - 2*se, UL = fit + 2*se)

```
### Plot fixed effects
```{r}
mean_gestation_birth <- format(round(mean(d$gestational_weeks_birth, na.rm = TRUE), 2), nsmall = 2)
sd_gestation_birth <- format(round(sd(d$gestational_weeks_birth, na.rm = TRUE), 2), nsmall = 2)

pred_mlm1 %>% 
  ggplot(aes(seg_week_cent_birth, fit)) +
  geom_ribbon(
    aes(ymin = LL, ymax = UL),
    fill = "grey80",
    alpha = .6
  ) +
  geom_line(size = 2) +
  geom_vline(
    aes(xintercept = 0),
    linetype = "dotted",
    size = 1
  ) +
  annotate(
    "text", 
    x = 15, 
    y = -.35,
    size = 5,
    colour = "grey45", 
    label = paste0("Mean[SD] gestational \nweeks = ", mean_gestation_birth , "[", sd_gestation_birth, "]")
  ) +
  geom_curve(
    aes(x = 11, y = -.5, xend = 0, yend = -.7), 
    arrow = arrow(length = unit(0.07, "inches")), 
    size = 0.2,
    colour = "grey65", 
    curvature = -.3
  ) +
  scale_y_continuous(breaks = seq.int(-1, 1, .25), limits = c(-1, 1)) +
  scale_x_continuous(breaks = seq.int(-40, 28, 8), limits = c(-40, 28)) +
  theme_pubr() +
  theme(
    strip.text = element_text(size = 16),
    axis.title.y = element_text(size = 20, face = "bold"),
    axis.title.x = element_text(size = 20, face = "bold"),
    axis.text.y = element_text(size = 16),
    axis.text.x = element_text(size = 16),
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.position = "none"
  ) +
  labs(
    x = "Weeks from childbirth",
    y = "log(HCC) pg/mg"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/hair_cortisol/plots/hcc_piecewise_average.png",
  width = 6,
  height = 5
)
```

### Plot with raw data
```{r}
pred_mlm1 %>% 
  ggplot(aes(seg_week_cent_birth, fit)) +
  geom_ribbon(
    aes(ymin = LL, ymax = UL),
    fill = "grey80",
    alpha = .6
  ) +
  geom_line(size = 2) +
  geom_point(
    data = d, 
    aes( y= hcc_log_m),
    size = .5, 
    alpha = .3
  ) +
  geom_line(
    data = d , 
    aes(y = hcc_log_m, group = ID),
    size=.2, 
    alpha=.3
  ) +
  scale_y_continuous(breaks = seq.int(-2, 5, .5)) +
  scale_x_continuous(breaks = seq.int(-40, 28, 8), limits = c(-40, 28)) +
  theme_pubr() +
  theme(
    strip.text = element_text(size = 16),
    axis.title.y = element_text(size = 20, face = "bold"),
    axis.title.x = element_text(size = 20, face = "bold"),
    axis.text.y = element_text(size = 16),
    axis.text.x = element_text(size = 16),
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.position = "none"
  ) +
  labs(
    x = "Weeks from childbirth",
    y = "log(HCC) pg/mg"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/hair_cortisol/plots/hcc_piecewise_id.png",
  width = 7,
  height = 5
)
```

## Examine effect of psychosocial adversity
```{r}
mlm2 <- lmer(
  scale(hcc_log_m) ~ 
    scale(seg_week_preg) +
    scale(seg_week_post) +
    scale(crisys_mu) +
    scale(crisys_pc) +
    (scale(seg_week_preg) + scale(seg_week_post) + scale(segment) | ID),
  data = d
)

summary(mlm2)
mlm2_parameters <- model_parameters(mlm2)
mlm2_parameters
performance(mlm2)

```

## Visualize effect of psychosocial adversity

### Extract predicted values and SEs

```{r}
# asymptotic covariance matrix (variances and covariances of fixed effects)
acov_mlm2 <- vcov(mlm2) 

pred_mlm2 <- expand.grid(
  crisys_mu = 0,
  crisys_pc = seq(-3, 3, 1)
) %>% 
  mutate(
    seg_week_preg = 0,
    seg_week_post = 0
  ) 

xmat_mlm2 <-
  pred_mlm2 %>% 
  mutate(
    intercept = 1
  ) %>% 
  dplyr::select(
    intercept,
    seg_week_preg,
    seg_week_post,
    crisys_mu,
    crisys_pc
  ) %>% 
  as.matrix()

# predicted values
fixed_mlm2 <- summary(mlm2)$coefficients[,"Estimate"]

pred_mlm2$fit <- 
  xmat_mlm2 %*% fixed_mlm2

# standard errors
pred_mlm2$se <- 
  xmat_mlm2 %*% acov_mlm2 %*% t(xmat_mlm2) %>% 
  diag() %>% 
  sqrt()

# CIs
pred_mlm2 <- 
  pred_mlm2 %>% 
  mutate(LL = fit - 2*se, UL = fit + 2*se) 
```

### Plot fixed effect
```{r}
pred_mlm2 %>% 
  ggplot(aes(crisys_pc, fit)) +
  geom_ribbon(
    aes(ymin = LL, ymax = UL),
    fill = "grey80",
    alpha = .6
  ) +
  geom_line(size = 2) +
  scale_y_continuous(breaks = seq.int(-1, 1, .25), limits = c(-1, 1)) +
  scale_x_continuous(breaks = seq.int(-3, 3, 1)) +
  theme_pubr() +
  theme(
    strip.text = element_text(size = 16),
    axis.title.y = element_text(size = 20, face = "bold"),
    axis.title.x = element_text(size = 20, face = "bold"),
    axis.text.y = element_text(size = 16),
    axis.text.x = element_text(size = 16),
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.position = "none"
  ) +
  labs(
    x = "Within-individual change\nin psychosocial adversity",
    y = "log(HCC) pg/mg"
  )

ggsave(
  "~/Box/lucy_king_files/BABIES/hair_cortisol/plots/hcc_crisyspc.png",
  width = 6,
  height = 5
)
```


## Examine effect of despressive symptoms
```{r}
mlm3 <- lmer(
  hcc_log_m ~ 
    scale(seg_week_preg) +
    scale(seg_week_post) +
    scale(cesd_mu) +
    scale(cesd_pc) +
    (scale(seg_week_preg) + scale(seg_week_post) + scale(segment) | ID),
  data = d
)

summary(mlm3)
mlm3_parameters <- model_parameters(mlm3)
mlm3_parameters
performance(mlm3)
```


## Examine covariates

### Age
```{r}
mlm4 <- lmer(
  scale(hcc_log_m) ~ 
    scale(seg_week_preg) +
    scale(seg_week_post) +
    scale(mom_age_t1) +
    (scale(seg_week_preg) + scale(seg_week_post) + scale(segment) | ID),
  data = d
)

summary(mlm4)
mlm4_parameters <- model_parameters(mlm4)
mlm4_parameters
performance(mlm4)

```


### Season
```{r}
# Fall is baseline
contrasts(d$season) = cbind(
  "SPvF" = c(-1/4, 1-(1/4), -1/4, -1/4),
  "SUvF" = c(-1/4, -1/4, 1-(1/4), -1/4),
  "WvF" = c(-1/4, -1/4, -1/4, 1-(1/4))
)

mlm5_fall <- lmer(
  scale(hcc_log_m) ~ 
    scale(seg_week_preg) +
    scale(seg_week_post) +
    season +
    (scale(seg_week_preg) + scale(seg_week_post) + scale(segment) | ID),
  data = d
)

anova(mlm5_fall)
summary(mlm5_fall)
mlm5_fall_parameters <- model_parameters(mlm5_fall)
mlm5_fall_parameters
performance(mlm5_fall)
```

### Estimated marginal means across seasons
```{r}
emmeans_season <- emmeans(mlm5_fall, specs = "season")
emmeans_season
```

### Hair chemical exposure
```{r}
contrasts(d$hair_chemexp_past5mo) = c(-.5, .5)

mlm6 <- lmer(
  scale(hcc_log_m) ~ 
    scale(seg_week_preg) +
    scale(seg_week_post) +
    hair_chemexp_past5mo +
    (scale(seg_week_preg) + scale(seg_week_post) + scale(segment) | ID),
  data = d
)

summary(mlm6)
mlm6_summer_parameters <- model_parameters(mlm6)
mlm6_summer_parameters
```

### Hair washing frequency
```{r}
mlm7 <- lmer(
  scale(hcc_log_m)~ 
    scale(seg_week_preg) +
    scale(seg_week_post) +
    scale(hair_wash_freq_mean) +
    (scale(seg_week_preg) + scale(seg_week_post) + scale(segment) | ID),
  data = d
)

summary(mlm7)
mlm7_summer_parameters <- model_parameters(mlm7)
mlm7_summer_parameters
```



